Computing the bicharacter of huge subspaces of polynomials

### 2019-07-23 article/article.tex
Review section on strategies to fit with context section. Move inert variables computations at the end and complete the subsection.

	The polynomial ring $R$ is endowed with an action of the symmetric group $\Sn$ by diagonal permutation of the columns of $X$ and of the general linear group $\GLr$ by diagonal action on the rows of $X$. 
	\begin{figure}[H]
		\centering
		\begin{tikzpicture}
		\draw node{$\begin{pmatrix}
			x_{1,1} & x_{1,2} & \cdots & x_{1,n} \\
			x_{2,1} & x_{2,2} & \cdots & x_{2,n} \\
			\vdots & \vdots & \ddots & \vdots \\
			x_{r,1} & x_{r,2} & \cdots & x_{r,n} \\
			\end{pmatrix}$ } (0,0);
		\draw (-2.4,.2)+(60:.25)
		[xscale=.5,yscale=3.2,->,rotate=80] arc(-40:290:.25);
		\node at (-3,0) {{\small$\GLr$}};
		\draw (-.5,-1.5)+(60:.25)
		[xscale=5.5,yscale=.6,->,rotate=-210] arc(-40:290:.25);
		\node at (0,-2) {{\small$\Sn$}};
		\end{tikzpicture}
		\caption{Simultaneous actions of $\GLr$ and $\Sn$ on $X$.}
	\end{figure}

@@

	Solving those differential equations would not be that tractable either. So instead we used operators, partial derivatives and polarization operators defined in the previous section. \\

@@

	The exact same strategy works to build the subspace generated by the Vandermonde $V_{\lambda}$ and its partial derivatives.

	Once a basis of $\harmonics_n^{(r)}$ or $\mathcal{E}_\lambda^{(r)}$ is constructed, there still remains to recover its bicharacter. For $\GLr$ that's given by the multigrading. For $\Sn$ some additional work has to be done (e.g. computing the traces of the action of permutations of all cycle types).

@@

	\begin{example} Compute $\mathcal{E}_{2,1}^{(2)}$ using the multidegree. 
		\begin{sagesilent}
			load("../compute_character.py")
		\end{sagesilent}
		First, we compute $\mathcal{E}_{2,1}^{(1)}$, for one row of variables. 
		\begin{sagecommandline}
			sage: v = vandermonde(Partition([2,1]))
			sage: v
			sage: derivatives = partial_derivatives(v.parent())
			sage: derivatives
			sage: E_21 = Subspace({v.multidegree(): [v]}, derivatives)
			sage: E_21.basis()
			sage: E_21.dimension()
		\end{sagecommandline}
		Then, we use polarization operators to introduce a second set of variables, and thus obtain $\mathcal{E}_{2,1}^{(2)}$, still with multidegree. 
		\begin{sagecommandline}
			sage: polarizators = polarization_operators(2)
			sage: polarizators
			sage: E_21 = PolarizedSpace(E_21, polarizators)
			sage: E_21.basis()
			sage: E_21.dimension()
			sage: E_21.dimensions()
		\end{sagecommandline}
		Compared with example \ref{H_3} in which all basis elements are grouped in one list with key $0$, here basis elements are classified by multidegree with the corresponding key. Operators are also separated by multidegree. Moreover, we have access to dimension by multidegree. 
	\end{example}

@@

	We first define two elements of the symmetric group algebra $\CC \Sn$. Note that we use the cyclic notation for permutations.
	Let $P_{\mu}$ be the subgroup of $\Sn$ of all row preserving permutations of $\mu$ and $Q_{\mu}$ the subgroup of the column preserving permutations of $\mu$. To determine the row and column preserving permutations, we consider the canonical Young tableau of $\mu$, that is to say the tableau filled in by integers from $1$ to $n$ in increasing order, from left to right and from bottom to top.

@@

	\begin{example} Consider $\Sym{3}$ and let $\mu = 2,1$.
		The canonical tableau associated is \tableau(12,3).\\
		Then $P_\lambda = \{1, (12)\}$ and $Q_\lambda = \{1, (13)\}$.

		For $\mu=2,1$, we have
		$a_{2,1} = 1 + (12)$ and
		$b_{2,1} = 1 - (13)$.

@@

	The proof of the above theorem is given by Fulton and Harris in \cite{FultonHarris1991}. 
	This theorem guarantees that $c_{\lambda}$ is an isotypic projector and that applying it on an adapted basis of $\mathcal{E}_{\lambda}$ gives representatives of the isotypic components of $\Sn$. \pauline{Need more details here.}  \\

	The action of $\GLr$ commutes with the action of $\Sn$, then the polarization operators commute with the Young symmetrizer. Therefore, we can compute independently each $\Sn$ isotypic component of $\mathcal{E}_\lambda^{(r)}$ by starting from a basis of $\mathcal{E}_\lambda^{(1)}$ that itself is adapted to its decomposition into isotypic components:
	$$\mathcal{E}_\lambda^{(1)}=:\bigoplus_\mu \mathcal{E}_{\lambda,\mu}^{(1)}.$$

	Then we use the polarization operators to obtain all the isotypic components of $\mathcal{E}_\lambda^{(r)}$ ($r>1$) from those of $\mathcal{E}_\lambda^{(1)}$ .

	\begin{example} \label{multigrading}
		Compute $\harmonics_3^{(2)} = \mathcal{E}_3^{(2)}$ using the Vandermonde determinant associated to $\lambda=3$. We still use the multigrading. 
		\begin{sagecommandline}
			sage: v = vandermonde(Partition([3]))
			sage: v
			sage: derivatives = partial_derivatives(v.parent())
			sage: E_3 = Subspace({v.multidegree(): [v]}, derivatives)
		\end{sagecommandline}
		At this stage, we project onto the isotypic components of $\Sym{3}$.
		\begin{sagecommandline}
			sage: E_3 = IsotypicComponent(E_3, 3)
			sage: E_3
		\end{sagecommandline}
		The result is now a dictionnary of subspaces, one for each isotypic component.
		\begin{sagecommandline}
			sage: E_3[Partition([3])].basis()
			sage: E_3[Partition([2,1])].basis()
			sage: E_3[Partition([1,1,1])].basis()
		\end{sagecommandline}
		We polarize each isotypic component. Note that, \texttt{PolarizedSpace} can receive a dictionary of subspaces, each subspace is then considered separately. 
		\begin{sagecommandline}
			sage: E_3 = PolarizedSpace(E_3, polarization_operators(2))
			sage: E_3[Partition([2,1])].basis()
		\end{sagecommandline}
	We print only the polarized isotypic component of type $2,1$ to avoid unreadable outputs in this example. 	
	\end{example}

@@

	Remark: For a fixed tableau $P$, denote by $\harmonics_P^{(r)}$ the space spanned by the higher Specht polynomials $(G_{P,Q})_Q$ under polarization. It's exactly the image of $\harmonics_n$ under the Young idempotent associated to $P$. For a given shape $\mu$, the $\harmonics_P^{(r)}$ are all conjugate under $\Sn$, and their direct sum is the isotypic component $\harmonics_{\mu}^{(r)}$.

@@

@@

	One thing that can slow down the computation is that the polynomials we are working with tend to be big. For example, for $\lambda \vdash 6$, the Vandermonde determinant $V_\lambda$, has $6!=720$ monomials. This means that we need a lot of space to store them and it takes time when we want to apply operators.

	However, since we use the Young symmetrizer to construct $\mathcal{E}_\lambda$, for all $v \in \mathcal{E}_\lambda$, $v$ is anti-symmetric with respect to the column stabilizer $Q_\mu$, with $\mu$ the type of the isotypic component of $v$. Indeed, when projecting a polynomial $p$ on its isotypic component using the Young idempotent $c_\mu$, by definition $p.a_\mu$ is symmetric with respect to the row preserving permutations of $\mu$ and then $v = p.a_\mu.b_\mu$ is anti-symmetric with respect to the column preserving permutations of $\mu$. 
	It's therefore sufficient to keep one representative monomial per $S_P$ orbit \pauline{$S_P$ orbit?} (and do some appropriate straightening whenever a new monomial is created). We chose to keep the smallest representative monomial in lexicographic order and call the result the \emph{normal reduced form} of $v$. \\

	Thus keeping reduced form of the polynomials decreases space complexity, and also time complexity when applying operators. Note that each time a polarization operator is applied to an element, the normal reduced form of the new element obtained is computed again.

@@

		$\Delta_{21} = -x_{2} \theta_{1} + x_{3} \theta_{1} + x_{1} \theta_{2} - x_{3} \theta_{2} - x_{1} \theta_{3} + x_{2} \theta_{3}$ is type $1,1,1$ and has reduced form $x_1\theta_2$.

		$\partial_{x_2}\Delta_{21} = -\theta_1 + \theta_3$ is type $2,1$ and has reduced form $-\theta_1$.
	\end{example}

	In practice, the antisymmetries are exploited first when projecting $\mathcal{E}_\lambda^{(1)}$ onto isotypic components. When applying the Young idempotent, the reduced form of each element is computed. From this step, the polynomials lie in a diagonal polynomial ring with antisymmetries, thus each polynomial carries the information of its antisymmetries.  
	Then, we maintain the elements of the bases of the isotypic components in their reduced form during polarization.

	\begin{example}
			Compute $\harmonics_3^{(2)} = \mathcal{E}_3^{(2)}$ exploiting antisymmetries. 
		\begin{sagecommandline}
			sage: v = vandermonde(Partition([3]))
			sage: derivatives = partial_derivatives(v.parent())
			sage: E_3 = Subspace({v.multidegree(): [v]}, derivatives)
		\end{sagecommandline}
		At this stage, we project onto the isotypic components of $\Sym{3}$ exploiting antisymmetries. One can check the difference between this computation and the one in example \ref{multigrading}. 
		\begin{sagecommandline}
			sage: E_3 = IsotypicComponent(E_3, 3, use_antisymmetry=True)
			%sage: E_3[Partition([3])].basis()
			sage: E_3[Partition([2,1])].basis()
			sage: E_3[Partition([1,1,1])].basis()
		\end{sagecommandline}
		We see that the parent of the polynomials carries the information about the antisymmetries to keep polynomials in normal reduced form.
		\begin{sagecommandline}
			sage: p = E_3[Partition([2,1])].basis().values()[0][0]
			sage: p
			sage: p.parent()
		\end{sagecommandline}
		Now we polarize each isotypic component still exploiting antisymmetries.  
		\begin{sagecommandline}
			sage: E_3 = PolarizedSpace(E_3, polarization_operators(2))
			sage: E_3[Partition([2,1])].basis()
		\end{sagecommandline}
	\end{example}

@@

	The space $\mathcal{E}_\lambda$ is symmetric with respect to permutation of rows of $X$, and such permutations exchange multigraded components. For example, for $r=3$, the homogeneous components of respective multidegree $(4,2,1)$ and $(2,1,4)$ coincide up to cycling of the rows. This means that there are polynomials in different homogeneous component that are the same up to a permutation of the sets of variables and that we have redundant information.

@@

	To achieve this, polarization operators can be tweaked. Whenever the result of a polarization is not of decreasing multidegree, straighten it by applying an appropriate permutation of the rows. It can then easily be shown that those tweaked polarization operators properly generate the whole $\mathcal{E}_\lambda^{(r)}$ space from $\mathcal{E}_\lambda^ {(1)}$. \pauline{Tweaked polarizators don't generate the whole space only the decreasing multidegree components.}

@@

		The polynomial $x_1^2 - 2x_1x_2 + 2x_2x_3 - x_3^2 \in \mathcal{H}_{3}$ has multidegree $(1,0)$ so the same polynomial with multidegree $(0,1)$ that is $y_1^2 - 2y_1y_2 + 2y_2y_3 - y_3^2$ also appears in $\mathcal{H}_{3}$ but it is not necessary to compute it explicitly.  
		%$-2x_{1} y_{1} y_{2} + x_{1} y_{2}^{2} \in \mathcal{H}_{3}$ so $-2y_{1} x_{1} x_{2} + y_{1} x_{2}^{2} \in \mathcal{H}_{3}$
	\end{example}

	In practice, when we polarize $\mathcal{E}_\lambda^{(1)}$, tweaked polarization operators are given in argument to \texttt{PolarizedSpace} instead of the classical ones. As said above, when specifying the row symmetries, only components of decreasing multidegree appear in the final result.

	\begin{example} Exploiting both antisymmetries of columns and row symmetries, the homogeneous components for $\mathcal{E}_3^{(2)}$ become the following.
		\begin{sagesilent}
			v = vandermonde(Partition([3]))
			derivatives = partial_derivatives(v.parent())
			E_3 = Subspace({v.multidegree(): [v]}, derivatives)
		\end{sagesilent}
		\begin{sagecommandline}
			sage: E_3 = IsotypicComponent(E_3, 3, use_antisymmetry=True)
			sage: E_3 = PolarizedSpace(E_3, polarization_operators(2, row_symmetry="permutation"))
			sage: E_3[Partition([1,1,1])].basis()
			sage: E_3[Partition([2,1])].basis()
			sage: E_3[Partition([3])].basis()
		\end{sagecommandline}	
	We observe that only decreasing multidegree components still appear, we don't have components of multidegree $(1,2)$, $(0,3)$, $(0,2)$ and $(0,1)$.
	\end{example}

@@

	As said before, the $\GLr$-character of $\mathcal{E}_\lambda$ is given by the multidegree of the homogeneous components. In other words, for a given isotypic component $\mathcal{E}_{\lambda,\mu}$, the $\GLr$-character associated to this component is given by $\sum_{\mathcal{E}_{\lambda,\mu}^{\langle d \rangle}} q^{d}$ where the sum is over all homogeneous component $\mathcal{E}_{\lambda,\mu}^{\langle d \rangle}$ of type $\mu$ and multidegree $d$ and $q^d = q_1^{d_1}q_2^{d_2}\dots$.

	When exploiting the row symmetries, since we don't compute only the homogeneous components of decreasing multidegree, we can not compute the character of $\GLr$ using such a summation. However, we know that if we have a component of multidegree $d=(d_1, d_2, \dots, d_k)$ , this correspond to a term $\sum_{\sigma \in \Sym{k}} q^{\sigma(d)} = m_d$ in the previous sum, thus we use monomials symmetric functions $m_d$ instead of summations to recover the character of $\GLr$ in each isotypic component.
	\pauline{Add time computation comparison (with and without optimisations) ?} \\

	For now we don't avoid compute some unnecessary homogeneous component since all we do is straighten the polynomials if they don't have decreasing multidegree. It would be even better if we could avoid completely the computation of those polynomials. It is not possible for now because we don't know yet which operators should not been applied to avoid those polynomials.

@@

	\pauline{I changed nothing in the two following subsections.}
	Building on the previous strategy, it would be tempting to only compute the subspace of $sl_r$ highest weight vectors within each $\harmonics_k^P$.

@@

	\subsection{Inert variables}
	\nicolas{2-3 paragraphs of announcement in the introduction}
	In their article \cite{GarsiaHaiman1993}, Garsia and Haiman also introduced a generalized version of the Vandermonde determinant and a generalization of the diagonal harmonics for two sets of variables called the Gasia-Haiman modules.

@@

	With a similar idea, François Bergeron recently introduced another generalization of the diagonal harmonics, with \emph{inert variables}. We add a new row $\bm{\theta} = (\theta_1, \theta_2, \dots, \theta_n)$ of variables in $X$ and call them inert. The symmetric group still acts by permutation of the columns of $X$, but $GL_r$ now acts diagonally only on the $r$ rows containing non-inert variables.

	Let $\lambda$ be a partition of $n$ and note $(a,b)$ the Cartesian coordinates of its cells. Then the determinant $V_{\lambda}$ associated to $\lambda$ is given by $$V_{\lambda} = \det (x_i^a \theta_i^b)_{1 \leq i \leq n \atop (a,b) \in \lambda} .$$

	\begin{example}
		The partition $2,1$ contains the cells $(0,0)$ and $(1,0)$ in the first row, and the cell $(0,1)$ in the second row, and $V_{2,1} = -x_2\theta_1 + x_3\theta_1 + x_1\theta_2 - x_3\theta_2 - x_1\theta_3 + x_2\theta_3$. 
	\end{example}

	Like before, we consider the smallest subspace containing $V_{\lambda}$ and closed by partial derivation and polarization, except that this time the inert variables do not intervene in these operators. We denote this space by $\mathcal{E}_{\lambda}^{(r)}$. The space of diagonal harmonics $\harmonics_n^{(r)}$ is thus a special case of $\mathcal{E}_{\lambda}^{(r)}$ for $\lambda = n$. \\

@@

	Like for $\harmonics_n^{(r)}$, the $\GLr \times \Sn$-character of $\mathcal{E}_\lambda^{(r)}$ is not easy to compute. \\

>>>>

	The polynomial ring $R$ is endowed with an action of the symmetric group $\Sn$ by diagonal permutation of the columns of $X$ and of the general linear group $\GLr$ by diagonal action on the rows of $X$. 
	\begin{figure}[H]
		\centering
		\begin{tikzpicture}
		\draw node{$\begin{pmatrix}
			x_{1,1} & x_{1,2} & \cdots & x_{1,n} \\
			x_{2,1} & x_{2,2} & \cdots & x_{2,n} \\
			\vdots & \vdots & \ddots & \vdots \\
			x_{r,1} & x_{r,2} & \cdots & x_{r,n} \\
			\end{pmatrix}$ } (0,0);
		\draw (-2.4,.3)+(60:.25)
		[xscale=.5,yscale=3.4,->,rotate=80] arc(-40:290:.25);
		\node at (-3,0) {{\small$\GLr$}};
		\draw (-.6,-1.5)+(60:.25)
		[xscale=5.8,yscale=.6,->,rotate=-210] arc(-40:290:.25);
		\node at (0,-2) {{\small$\Sn$}};
		\end{tikzpicture}
		\caption{Simultaneous actions of $\GLr$ and $\Sn$ on $X$.}
	\end{figure}

@@

	Solving those differential equations would not be tractable either. So instead we used operators, partial derivatives and polarization operators defined in the previous section. \\

@@

	Once a basis of $\harmonics_n^{(r)}$ is constructed, there still remains to recover its bicharacter. For $\GLr$ that's given by the multigrading. For $\Sn$ some additional work has to be done (e.g. computing the traces of the action of permutations of all cycle types).

@@

	\begin{example} Compute $\harmonics_{3}^{(2)}$ using the multidegree. 
		\begin{sagesilent}
			load("../compute_character.py")
		\end{sagesilent}
		First, we compute $\harmonics_{3}^{(1)}$, for one row of variables. 
		\begin{sagecommandline}
			sage: v = vandermonde(3)
			sage: v
			sage: derivatives = partial_derivatives(v.parent())
			sage: derivatives
			sage: H_3 = Subspace({v.multidegree(): [v]}, derivatives)
			sage: H_3.basis()
			sage: H_3.dimension()
		\end{sagecommandline}
		Then, we use polarization operators to introduce a second set of variables, and thus obtain $\harmonics_3^{(2)}$, still with multidegree. 
		\begin{sagecommandline}
			sage: polarizators = polarization_operators(2)
			sage: polarizators
			sage: H_3 = PolarizedSpace(H_3, polarizators)
			%sage: H_3.basis()
			sage: H_3.dimension()
			sage: H_3.dimensions()
		\end{sagecommandline}
		Compared with example \ref{H_3} in which all basis elements are grouped in one list with key $0$, here basis elements are classified by multidegree with the corresponding key. Operators are also separated by multidegree. Moreover, we have access to dimension by multidegree. 
	\end{example}

@@

	We first define two elements of the symmetric group algebra $\CC \Sn$. Note that we use the cyclic notation for permutations.
	Let $P_{\mu}$ be the subgroup of $\Sn$ of all row preserving permutations of $\mu$ and $Q_{\mu}$ the subgroup of all column preserving permutations of $\mu$.

	To determine the row and column preserving permutations, we consider the canonical Young tableau of $\mu$, that is to say the tableau filled in by integers from $1$ to $n$ in increasing order, from left to right and from bottom to top. Then a permutation is said row (resp. column) preserving if when applied to the canonical tableau of $\mu$, the entries of each row (resp. column) remains in the same row (resp. column) maybe in a different order (see example \ref{young_idempotent}).

@@

	\begin{example}\label{young_idempotent} Let $\mu = (2,1)$.
		The canonical tableau associated is \tableau(12,3).

		Then $P_\lambda = \{1, (12)\}$ and $Q_\lambda = \{1, (13)\}$.

		For $\mu=(2,1)$, we have
		$a_{2,1} = 1 + (12)$ and
		$b_{2,1} = 1 - (13)$.

@@

	The proof of the above theorem is given by Fulton and Harris in \cite{FultonHarris1991}. 
	This theorem guarantees that $c_{\lambda}$ is an isotypic projector and that applying it on an adapted basis of $\harmonics_{n}^{(r)}$ gives representatives of the isotypic components of $\Sn$. \pauline{Need more details here.}  \\

	The action of $\GLr$ commutes with the action of $\Sn$, then the polarization operators commute with the Young symmetrizer. Therefore, we can compute independently each $\Sn$ isotypic component of $\harmonics_{n}^{(r)}$ by starting from a basis of $\harmonics_{n}^{(1)}$ that itself is adapted to its decomposition into isotypic components:
	$$\harmonics_{n}^{(1)}=:\bigoplus_{\mu \vdash n} \harmonics_{\mu}^{(1)},$$

	where $\harmonics_{\mu}^{(1)}$ is the isotypic component of type $\mu$ of $\harmonics_{n}^{(1)}$.

	Then we use the polarization operators to obtain all the isotypic components $\harmonics_{\mu}^{(r)}$ of $\harmonics_n^{(r)}$ ($r>1$) from those of $\harmonics_n^{(1)}$ .

	\begin{example} \label{multigrading}
		Compute $\harmonics_3^{(2)}$ using the Vandermonde determinant for $n=3$. We still use the multigrading. 
		\begin{sagecommandline}
			sage: v = vandermonde(3)
			sage: derivatives = partial_derivatives(v.parent())
			sage: H_3 = Subspace({v.multidegree(): [v]}, derivatives)
		\end{sagecommandline}
		At this stage, we project onto the isotypic components of $\Sym{3}$.
		\begin{sagecommandline}
			sage: H_3 = IsotypicComponent(H_3, 3)
			sage: H_3
		\end{sagecommandline}
		The result is now a dictionnary of subspaces, one for each isotypic component.
		\begin{sagecommandline}
			sage: H_3[Partition([3])].basis()
			sage: H_3[Partition([2,1])].basis()
			sage: H_3[Partition([1,1,1])].basis()
		\end{sagecommandline}
		We polarize each isotypic component. Note that, \texttt{PolarizedSpace} can receive a dictionary of subspaces, each subspace is then considered separately. 
		\begin{sagecommandline}
			sage: H_3 = PolarizedSpace(H_3, polarization_operators(2))
			sage: H_3[Partition([2,1])].basis()
		\end{sagecommandline}
	We print only the polarized isotypic component of type $(2,1)$ to avoid unreadable outputs in this example. 	
	\end{example}

@@

	For a fixed tableau $P$, denote by $\harmonics_P^{(r)}$ the space spanned by the higher Specht polynomials $(G_{P,Q})_Q$ under polarization. It's exactly the image of $\harmonics_n$ under the Young idempotent associated to $P$. For a given shape $\mu$, the $\harmonics_P^{(r)}$ are all conjugate under $\Sn$, and their direct sum is the isotypic component $\harmonics_{\mu}^{(r)}$.

@@

	However, for now we have not managed to fully implement this strategy but are still working on it.

@@

	One thing that can slow down the computation is that the polynomials we are working with tend to be big. For example, for $\lambda \vdash 6$, the Vandermonde determinant $V_\lambda$, is a homogeneous polynomial of degree $\sum_{i=1}^{5}i = 15$ with $6!=720$ monomials . This means that we need a lot of space to store them and it takes time when we want to apply operators.

	However, since we use the Young symmetrizer to construct $\harmonics_{n}^{(r)}$, for all $v \in \harmonics_{n}^{(r)}$, $v$ is anti-symmetric with respect to the column stabilizer $Q_\mu$, with $\mu$ the type of the isotypic component of $v$. Indeed, when projecting a polynomial $p$ on its isotypic component using the Young idempotent $c_\mu$, by definition $p.a_\mu$ is symmetric with respect to the row preserving permutations of $\mu$ and then $v = p.a_\mu.b_\mu$ is anti-symmetric with respect to the column preserving permutations of $\mu$. 
	It's therefore sufficient to keep one representative monomial per $S_P$ orbit \pauline{$S_P$ orbit?} (and do some appropriate straightening whenever a new monomial is created). We chose to keep the smallest representative monomial in lexicographic order and call the result the \emph{normal reduced form} of $v$. \\

	Thus keeping reduced forms of the polynomials decreases space complexity, and also time complexity when applying operators. Note that each time a polarization operator is applied to an element, the normal reduced form of the new element obtained is computed again.

@@

		$V_{3} = -x_{1}^{2} x_{2} + x_{1} x_{2}^{2} + x_{1}^{2} x_{3} - x_{2}^{2} x_{3} - x_{1} x_{3}^{2} + x_{2} x_{3}^{2}$ is of type $(1,1,1)$ and has reduced form $x_1^2x_2$.

		$\partial_{x_2}V_{3} = -x_{1}^{2} + 2 x_{1} x_{2} - 2 x_{2} x_{3} + x_{3}^{2}$ is of type $(2,1)$ and has reduced form $x_1^2 - 2x_1x_2$.
	\end{example}

	In practice, the antisymmetries are exploited first when projecting $\harmonics_n^{(1)}$ onto its isotypic components. When applying the Young idempotent, the reduced form of each element is computed. From this step, the polynomials lie in a diagonal polynomial ring with antisymmetries, thus each polynomial carries the information of its antisymmetries.  
	Then, we maintain the elements of the bases of the isotypic components in their reduced form during polarization.

	\begin{example}
			Compute $\harmonics_3^{(2)}$ exploiting antisymmetries. 
		\begin{sagecommandline}
			sage: v = vandermonde(Partition([3]))
			sage: derivatives = partial_derivatives(v.parent())
			sage: H_3 = Subspace({v.multidegree(): [v]}, derivatives)
		\end{sagecommandline}
		At this stage, we project onto the isotypic components of $\Sym{3}$ exploiting antisymmetries. One can check the difference between this computation and the one in example \ref{multigrading}. 
		\begin{sagecommandline}
			sage: H_3 = IsotypicComponent(H_3, 3, use_antisymmetry=True)
			%sage: H_3[Partition([3])].basis()
			sage: H_3[Partition([2,1])].basis()
			sage: H_3[Partition([1,1,1])].basis()
		\end{sagecommandline}
		We see that the parent of the polynomials carries the information about the antisymmetries to keep polynomials in normal reduced form.
		\begin{sagecommandline}
			sage: p = H_3[Partition([2,1])].basis().values()[0][0]
			sage: p
			sage: p.parent()
		\end{sagecommandline}
		Now we polarize each isotypic component still exploiting antisymmetries.  
		\begin{sagecommandline}
			sage: H_3 = PolarizedSpace(H_3, polarization_operators(2))
			sage: H_3[Partition([2,1])].basis()
		\end{sagecommandline}
	\end{example}

@@

	The space $\harmonics_{n}^{(r)}$ is symmetric with respect to permutation of rows of $X$, and such permutations exchange multigraded components. For example, for $r=3$, the homogeneous components of respective multidegree $(4,2,1)$ and $(2,1,4)$ coincide up to cycling of the rows. This means that there are polynomials in different homogeneous components that are the same up to a permutation of the sets of variables and that we have redundant information.

@@

	To achieve this, polarization operators can be tweaked. Whenever the result of a polarization is not of decreasing multidegree, straighten it by applying an appropriate permutation of the rows. It can then easily be shown that those tweaked polarization operators properly generate the whole $\harmonics_{n}^{(r)}$ space from $\harmonics_n^ {(1)}$. \pauline{Tweaked polarizators don't generate the whole space only the decreasing multidegree components, right?}

@@

		The polynomial $x_1^2 - 2x_1x_2 + 2x_2x_3 - x_3^2 \in \mathcal{H}_{3}^{(2)}$ has multidegree $(1,0)$ so the same polynomial with multidegree $(0,1)$ that is $y_1^2 - 2y_1y_2 + 2y_2y_3 - y_3^2$ also appears in $\mathcal{H}_{3}^{(2)}$ but it is not necessary to compute it explicitly.  
		%$-2x_{1} y_{1} y_{2} + x_{1} y_{2}^{2} \in \mathcal{H}_{3}$ so $-2y_{1} x_{1} x_{2} + y_{1} x_{2}^{2} \in \mathcal{H}_{3}$
	\end{example}

	In practice, when we polarize $\harmonics_{n}^{(1)}$, tweaked polarization operators are given in argument to \texttt{PolarizedSpace} instead of the classical ones. As said above, when specifying the row symmetries, only components of decreasing multidegree appear in the final result.

	\begin{example} Exploiting both antisymmetries of columns and row symmetries, the homogeneous components for $\harmonics_3^{(2)}$ become the following.
		\begin{sagesilent}
			v = vandermonde(Partition([3]))
			derivatives = partial_derivatives(v.parent())
			H_3 = Subspace({v.multidegree(): [v]}, derivatives)
		\end{sagesilent}
		\begin{sagecommandline}
			sage: H_3 = IsotypicComponent(H_3, 3, use_antisymmetry=True)
			sage: H_3 = PolarizedSpace(H_3, polarization_operators(2, row_symmetry="permutation"))
			sage: H_3[Partition([1,1,1])].basis()
			sage: H_3[Partition([2,1])].basis()
			sage: H_3[Partition([3])].basis()
		\end{sagecommandline}	
	We observe that only decreasing multidegree components still appear, we don't have components of multidegree $(1,2)$, $(0,3)$, $(0,2)$ and $(0,1)$.
	\end{example}

@@

	As said before, the $\GLr$-character of $\harmonics_n^{(r)}$ is given by the multidegree of the homogeneous components. In other words, for a given isotypic component $\harmonics_{\mu}^{(r)}$, $\mu \vdash n$, the $\GLr$-character associated to this component is given by $$\sum_{\harmonics_{\mu,d}^{(r)}} q^{d}$$ where the sum is over all homogeneous component $\harmonics_{\mu,d}^{(r)}$ of type $\mu$ and multidegree $d$ and $q^d = q_1^{d_1}q_2^{d_2}\dots$.

	When exploiting the row symmetries, since we compute only the homogeneous components of decreasing multidegree, we can not compute the character of $\GLr$ using such a summation. However, we know that if we have a component of multidegree $d=(d_1, d_2, \dots, d_k)$ , this correspond to a term $\sum_{\sigma \in \Sym{k}} q^{\sigma(d)} = m_d$ in the previous sum, thus we use monomials symmetric functions $m_d$ instead of summations to recover the character of $\GLr$ in each isotypic component.
	\pauline{Add time computation comparison (with and without optimisations) ?} \\

	For now we don't avoid compute some unnecessary homogeneous component since all we do is straighten the polynomials if they don't have decreasing multidegree. It would be even better if we could avoid completely the computation of those polynomials. It is not possible for now because we don't know yet which operators should not been applied to avoid those polynomials. \\

	Finding an efficient way to generate $\harmonics_{n}^{(r)}$ is the main part of the computation. Once we have the isotypic components separated by multidegree, we obtain immediatly the character.

	\begin{example} We compute the character of $\harmonics_{3}^{(r)}$. Since we exploit row symmetry, we have to specify it in order to use monomials instead of sums as explain above. 
		\begin{sagecommandline}
			sage: character(H_3, row_symmetry="permutation")
		\end{sagecommandline}
	We finally obtain that the character of $\harmonics_{3}^{(r)}$ for $r \geq 2$ is $$\harmonics_{3}^{(r)} = 1 \otimes s_{3} +(s_{1} + s_{2}) \otimes s_{2,1} +(s_{1,1} + s_{3}) \otimes s_{1,1,1}.$$
	\end{example}

@@

	\pauline{I changed nothing in the two following subsections. I leave it to you Nicolas.}
	Building on the previous strategy, it would be tempting to only compute the subspace of $sl_r$ highest weight vectors within each $\harmonics_k^P$.

@@

	\subsection{Inert variables}
	\nicolas{2-3 paragraphs of announcement in the introduction}
	In their article \cite{GarsiaHaiman1993}, Garsia and Haiman introduce a generalized version of the Vandermonde determinant and a generalization of the diagonal harmonics for two sets of variables called the Gasia-Haiman modules.

@@

	With a similar idea, François Bergeron introduce another generalization of the diagonal harmonics, with \emph{inert variables}. We add a new row $\bm{\theta} = (\theta_1, \theta_2, \dots, \theta_n)$ of variables in $X$ and call them inert. The symmetric group still acts by permutation of the columns of $X$, but $GL_r$ now acts diagonally only on the $r$ rows containing non-inert variables.

	\begin{figure}[H]
		\centering
		\begin{tikzpicture}
		\draw node{$\begin{pmatrix}
			x_{1,1} & x_{1,2} & \cdots & x_{1,n} \\
			x_{2,1} & x_{2,2} & \cdots & x_{2,n} \\
			\vdots & \vdots & \ddots & \vdots \\
			x_{r,1} & x_{r,2} & \cdots & x_{r,n} \\ 
			\theta_{1} & \theta_{2} & \cdots & \theta_{n}
			\end{pmatrix}$ } (0,0);
		\draw (-2.4,.6)+(60:.25)
		[xscale=.5,yscale=3.5,->,rotate=80] arc(-40:290:.25);
		\node at (-3,0.3) {{\small$\GLr$}};
		\draw (-.6,-1.7)+(60:.25)
		[xscale=5.8,yscale=.6,->,rotate=-210] arc(-40:290:.25);
		\node at (0,-2.1) {{\small$\Sn$}};
		\draw (-3,-.85) -- (-1.5,-.85);
		\end{tikzpicture}
		\caption{Simultaneous actions of $\GLr$ and $\Sn$ on $X$ with inert variables.}
	\end{figure}

	Let $\lambda$ be a partition of $n$ and note $(a,b)$ the Cartesian coordinates of its cells. Then the determinant $V_{\lambda}$ associated to $\lambda$ is given by $$V_{\lambda} := \det (x_i^a \theta_i^b)_{1 \leq i \leq n \atop (a,b) \in \lambda} .$$

	\begin{example}
		The partition $(2,1)$ contains the cells $(0,0)$ and $(1,0)$ in the first row, and the cell $(0,1)$ in the second row, and $V_{2,1} = -x_2\theta_1 + x_3\theta_1 + x_1\theta_2 - x_3\theta_2 - x_1\theta_3 + x_2\theta_3$. 
	\end{example}

	Like before, we consider the smallest subspace containing $V_{\lambda}$ and closed by partial derivation and polarization, except that this time the inert variables do not intervene in these operators. We denote this space by $\mathcal{E}_{\lambda}^{(r)}$. The space of diagonal harmonics $\harmonics_n^{(r)}$ is thus a special case of $\mathcal{E}_{\lambda}^{(r)}$ for $\lambda = (n)$. \\

@@

	Like for $\harmonics_n^{(r)}$, the $\GLr \times \Sn$-character of $\mathcal{E}_\lambda^{(r)}$ is not easy to compute but hopefully most of the strategies we used to compute the character of $\harmonics_{n}^{(r)}$ work for $\mathcal{E}_\lambda^{(r)}$.

	Therefore, we, once again, use polarization to compute $\mathcal{E}_\lambda^{(r)}$ from $\mathcal{E}_\lambda^{(1)}$, exploit the multigrading. We use the same isotypic projector to obtain the isotypic components, then we can still exploit the columns antisymmetries. 
%	\begin{example} ~
%		
%		$V_{21} = -x_{2} \theta_{1} + x_{3} \theta_{1} + x_{1} \theta_{2} - x_{3} \theta_{2} - x_{1} \theta_{3} + x_{2} \theta_{3}$ is of type $(1,1,1)$ and has reduced form $x_1\theta_2$.
%		
%		$\partial_{x_2}V_{21} = -\theta_1 + \theta_3$ is of type $(2,1)$ and has reduced form $-\theta_1$.
%	\end{example}
	Finally, $\mathcal{E}_\lambda^{(r)}$ is still symmetric with respect to a permutation of the rows of $X$ containing classical variables so we exploit row symmetries in this case too.\\

	\begin{example} In the following example, we use different strategies to compute $\mathcal{E}_{2,1}^{(2)}$.
